Supermassive black holes and their host spheroids \\ II. The red and blue sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram

### 2015-10-09 papers/MbhMsph/replytoreferee_II.txt

- The referee is correct, and we are aware of this. Unlike many papers, we believe that we have not under-
  estimated the uncertainties associated to the bulge best-fit parameters. It is possible that some of the 
  results taken from the literature are "grossly wrong", however the exclusion of extreme outliers would 
  affect only the sigma_3 error bars, i.e. the uncertainties associated to the "grade 3" fits, which constitute 
  only 14% of all fits. The exclusion of extreme outliers would have a very small impact on the sigma_1 and 
  sigma_2 errorbars (we learned this from Savorgnan et al. 2013, where we applied a very similar method). 
  In conclusion, the exclusion of extreme outliers would decrease the uncertainties associated to 14% of 
  all galaxies, with little or no effects on the measurement of the intrinsic scatter.

- The referee is correct in saying that the last point in our Conclusion Section is not a new result and,  
  indeed, we didn't mean to deliver this result as a novelty. The Introductions of Papers I and II present 
  some points about black hole mass scaling relations that are currently under debate. One of these points 
  is the level of intrinsic scatter in the black hole mass - spheroid/galaxy luminosity correlation. Some 
  studies (e.g. Erwin & Gadotti 2012, Beifiori et al. 2012, Kormendy & Ho 2013) claimed that the spheroid 
  luminosity is a better tracer of the black hole mass, but the most recent study by Lasker et al. reported 
  that the spheroid luminosity and the galaxy luminosity are equally good proxies for the black hole mass. 
  Our work solves this controversy by showing that the spheroid luminosity and the galaxy luminosity correlate 
  equally well with the black hole mass when only early-type galaxies are considered (Lasker's sample has only 
  4 spiral galaxies). However, when the galaxy sample contains a reasonable number of spiral galaxies, the 
  correlation with the spheroid luminosity has less intrinsic scatter than that with the galaxy luminosity. 
  At our best knowledge, this has never been shown before in the literature. 
  Additional references have been added to the last point in our Conclusion Section.

- We think that the disk of the Sombrero galaxy is an intermediate-scale one. 
  Our light profile has been truncated at 160'', but we have information about the isophotal parameters out to
  250'' and the ellipticity profile indicates the presence of an intermediate-scale disk that may be truncated 
  at 200''. 
  The bulge/disk decomposition of Gadotti & Sanchez-Janssen (figure 2) uses a more extended light profile 
  (out to 450'') and is consistent with ours. 
  Although we did our best at scrutinizing the available literature, we unfortunately missed this work, which
  has now been referenced in Paper I. 
  We note that the Sombrero galaxy has been classified as S0/Sp in Paper II, therefore it has not been included 
  in the early-type nor in the late-type bin. We consider the morphological classification of this galaxy quite 
  challenging, in the sense that the Sombrero has the structure of a E/S0 galaxy but it also has features typical 
  of a spiral galaxy (rings and dust at the edge of the disc; note that NED refers to the Sombrero as an Sa spiral
  galaxy), therefore we feel it's safer to exclude it from the morphological analysis of the Mbh-M diagram.

- We prefer not to reference Burstein et al. because many of their claimed discs were falsely identified 
  due to poor R^(1/4) decompositions.

@@

- A footnote has been added in Section 4.3 to remind the reader why a slope~1 is consistent with a dry-merging 
  formation scenario, although it simply repeats what already stated in the Introduction.

- A (B-V) galaxy color vs. log(galaxy stellar mass) plot is attached. The early-type galaxies are seen to occupy 
  the region of the red sequence, and the spiral galaxies are located in the region of the blue cloud. 
  The correspondence between galaxy morphological type (early vs. late) and position in the color-mass diagram 
  (red sequence vs. blue cloud) is a very well established result and we don't feel the need of showing this 
  (poor) diagram in the paper. 
  TBD ATTACH PLOT

- We fully agree with the referee on this point, as extensively discussed in Section 4.2.1. We don't think that 
  classical and pseudo-bulges can be distinguished based on their Sersic index. However, this has been common 
  practice in the recent literature, and here we are showing that n<2 bulges are not offset from the correlation
  traced by n>2 bulges. The referee might appreciate that we never referred to our n<2 bulges merely as pseudo-bulges, 
  but we always specified "claimed by some to be pseudo-bulges". In addition, in the text we have questioned the 
  n<2 criterion used by some to identify pseudo-bulges.

- We never claimed that the break of the black hole - bulge mass correlation is due to the bulge mass. 
  In the Introduction, we anticipated that, for consistency with other known scaling relations (Lsph-sigma and 
  Mbh-sigma), a bent Mbh-Lsph correlation is expected, with the bend occurring around Mbh~10^8. 
  In the Mbh-Lsph and Mbh-Msph diagrams we identified two different sequences (early vs. late) precisely according 
  to the morphology of galaxies. Indeed, the break is caused by the galaxy morphology. 
  We did not identify pseudo-bulges with n<2 bulges. 
  In Figure 5, the red dashed line marks the correlation for ellipticals+lenticulars (as stated in the caption). 
  The correlation for ellipticals only is not shown in this Figure. The spiral galaxies can be seen not following 
  the E+S0 correlation. 
  Concerns about the lack of a correlation for spiral galaxies are not correct, and have already been addressed in 
  the last paragraph of Section 4.2. The correlation for spiral galaxies has been measured and quantified with 3 
  different linear regression routines. If there was no correlation for spiral galaxies, the uncertainties associated 
  to the slope and intercept of the linear regression would be extremely large.  
  Addressing the gas content of the galaxies in the Mbh-M diagram is definitely a priority for our future research, 
  and we are planning to tackle this point with a dedicated study.

### 2015-10-09 papers/data_paper/replytoreferee_I.txt

- The tone of the title was not meant to be presumptuous, but we now understand that it is prone to misinterpretation, 
  therefore we have modified it. The new title reads: "Galaxy vivisection".

- We agree that some disk galaxies, like NGC 300, can be bulge-less. However, all galaxies in our sample are either 
  ellipticals/lenticulars or early-type spirals (Sa-Sc), with the only exception of NGC 4945 (Scd). 
  By definition, these galaxies have a spheroid/bulge component, which is what we modelled with a Sersic profile, 
  without attempting to distinguish between classical and pseudo-bulges. The text in Section 3.5 was actually misleading 
  and has been modified.
  Throughout the text, we used the terms "spheroid" and "bulge" as synonims, although we recognize that some bulges are 
  "disk-like" rather than "spheroidal-like". This point has now been clarified in Footnote 1.

- We did not attempt to fit down- or up-bendings of large-scale disks. We scrutinized the 
  available literature for all individual galaxies, and we paid particular attention to the papers that recorded disk breaks. 
  The surface brightness profiles of galaxies that present this feature were truncated well before the occurence of the 
  bending. This was not explicit in the text, and we now added a sentence at the end of Section 3.1.

- The 1D equivalent axis fit does account for the varying ellipticity because the equivalent 
  axis is calculated as Req = Rmaj * (1-ellipticity)^0.5. 
  M87 has a strong ellipticity gradient. The 2D model cannot account for this gradient, and this can be seen in the 2D 
  residual image which looks like a "butterfly pattern". 
  Ideally, galaxies should be fit in 3D, rather than 2D or 1D. 
  However, because the 1D fit tracks the changes in the observed P.A. and ellipticity, the geometric mean axis 
  (equivalent to a circularised axis) enables one to accurately recover the total and component magnitudes 
  in a way that the 2D model cannot.

- The text in Section 3.5 was not accurate, and we have now clarified this point. All bars have been fit with a Ferrer function, 
  except for NGC 4151 where we preferred to use a Sersic model. Disk-like components embedded in the bulges/spheroids have 
  been fit with a low-n Sersic profile and, in a few cases (when the Sersic model was degenerate with the other model-
  components), with a Ferrer function. 
  Gadotti & De Souza (2006) report that flat bars are typical of earlier type spirals, whereas exponential bars are common 
  in later type spirals. This is in agreement with our findings: our galaxy sample lacks of later-type spirals and we did 
  not detect any bar with an exponential profile.

- Sheth et al. (2010, the paper that presents the S4G collaboration) refers to Jun & Im (2008) to point out that the 3.6 um
  band is indeed a better tracer of the stellar mass than the K-band, which still suffers from dust absorption (Driver et al.
  2008, 678, L101).

- In Figure 7 we compared the 2D Sersic indices with the 1D equivalent-axis Sersic indices. These two quantities are not 
  expected to be exactly the same, but they should be approximatively similar. There are four most obvious outliers in this 
  plot: NGC 821, NGC 3377, NGC 4945, and NGC 5576. Their individual cases are discussed in the Appendix. 
  For M87, the 2D Sersic index (4.3) and the 1D equivalent-axis Sersic index (5.9) are in fairly good agreement. 
  The large discrepancy between the M87's 1D major-axis Sersic index and the other two measurements is due to the strong ellipticity
  gradient that this galaxy has. In fact, the 2D model cannot account for such gradient (whereas the 1D equivalent-axis fit can) 
  and the 2D residuals show a significant ``butterfly'' feature caused by this well-understood problem.

- The aim of this work is to investigate scaling relations between the black hole mass, Mbh, and several properties, X, of the host 
  spheroid/bulge. These correlations are commonly expressed as power laws: Mbh \propto X^a, i.e. log(Mbh) = a*log(X) + b. 
  Therefore we are interested in the logarithmic values of the Sersic indices and effective radii, along with their associated 
  uncertainties.

- The point that Footnote 2 was trying to make is the following. Sani et al., Vika et al. and Beifiori et al. did not find a 
  correlation between black hole mass and Sersic index. However, after averaging their Sersic indices together, a strong correlation
  was recovered. This is telling us that the measurements obtained by the individual authors were "noisy", i.e. non accurate. 
  Averaging the Sersic indices together reduces the noise and shows that the black hole mass is indeed correlated to the Sersic 
  index. However, we removed the second sentence of Footnote 2; we will come back to this point in a future paper.

- The claim about the anti-truncation of NGC 4459 has been clarified in the Appendix.

- Thanks for pointing out the missing reference to Erwin & Gadotti (2012), which has now been added.

- The Spitzer cookbook does not explain why the overlap correction is not recommended for 3.6 um images, it just states so. 
  It might be that the cookbook does not consider mosaicking images belonging to different "Astronomical Observation Requests", 
  for which we found that the overlap correction was needed to remove the chessboard effect.

- We did check that the PSF does not significantly vary across the various mosaics. In each mosaic, we selected up to 10 bright 
  non-saturated stars and we characterized them with a Moffat function (using IRAF imexamine). For each mosaic, we averaged 
  together the best-fit parameters of the stars and obtained an average FWHM and beta (the PSF parameters specific of each mosaic). 
  The variation of the Moffat FWHM across all mosaics is +-0''.1; the variation of the Moffat beta is +-2.0.
  The use of a universal PSF is justified for the following reasons: 
  i) the PSF-convolution is way more sensitive to the value of the FWHM than beta, therefore having a 50% variation in beta is 
  not an issue as long as the variation in FWHM is small;
  ii) the use of a non-signal-to-noise-weighted fitting scheme minimizes biases from a non-accurate PSF description;
  iii) not all mosaics have enough stars suitable for PSF characterization; rather than having an individual PSF for each mosaic 
  (which would have been based on only 1-2 stars for some mosaics), we preferred using a universal PSF;
  iv) the IRAC instrument handbook guarantees a pretty stable PSF across time.

- A reference to Gadotti (2008) has been added in Section 3.5.

- Regarding the comment on weighting schemes in Section 3.3, we attach two simple, but hopefully illustrative figures. 
  In both figures, the left hand panel shows a perfect R^(1/4) profile.
  In one figure no weighting scheme is used, while in the other a signal-to-noise-weighting scheme is used. 
  The results of the fit are the same. 
  Now in the right hand-panel, exactly the same partially depleted core has been introduced in both figures.  
  When the R^(1/4) model is fit with the different weighting schemes, the signal-to-noise-weighted scheme can be seen to perform worse.  
  In fact, any central deviation from the assumed model (in this case an R^(1/4) model, but we could have used a Sersic model) 
  results in the signal-to-noise-weighted fitting scheme performing worse.  
  Such deviations arise from: additional star clusters, dust, depleted cores, AGN, non accurate PSF, etc.  
  Unless one correctly accounts for all of these things (which realistically doesn't happen), then the signal-to-noise-weighted scheme 
  performs worse than a scheme with no weighting when it comes to matching the global light profile. 
  ATTACH FIGURES TBD

- References to Sheth et al. and Kim et al. have been added.

- The contrast in Figure 5 was indeed not appropriate and has been adjusted.

- Section 4.2.1, item 2: the typo has been corrected.

- Section 4.2.2: a sentence has been added to explain the different sigma symbols.

- The edge-on disc of NGC 1332 is clearly identified in the galaxy image. The ellipticity profile of the galaxy isophotes indicates 
  that the disc is intermediate-scale. We modified the caption of Figure 35 to better illustrate this point.

- The fourth paragraph of Section 3.5 explains how we decided to include a nuclear component in our models.

- We tried to model the bar of NGC 4945 with a Ferrer function or a Sersic model, but we did not obtain a satisfactory result. 
  We now realize that NGC 4945 is the only later type (Scd) spiral galaxy in our sample, and its bar might might not be well described 
  by a Ferrer function. However, a Sersic model did not provide a good fit of the bar either. It is likely that the edge-on 
  inclination of the galaxy complicated the extraction of the surface brightness profile with ELLIPSE. 
  Our best solution is still to exclude the data in the range where the bar dominates the light and model only the bulge and the 
  large-scale disk. We note that we classified the fit of this galaxy as "grade 2".

- Kormendy & Ho (2013) did not perform bulge/disk decompositions for their galaxies, but they collected results from various sources 
  in the literature. We wanted to compare the results from our decompositions with the results from other systematic studies that 
  performed galaxy modelling in a consistent way.

### 2015-10-09 papers/MbhMsph/mm.tex
friday and i love u kri

%\title{Supermassive black hole and host bulge affairs \\ II. The early and late sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram}
\title{Supermassive black hole and host spheroid affairs \\ II. The red and blue sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram}

@@

Kinematical information \citep{atlas3dIII-MNRAS,scott2014,arnold2014} was used 

@@

errors are dominated by systematics rather than Poisson noise.} 

@@

%More than half of all morphologically classified elliptical galaxies in the ATLAS$^{\rm 3D}$ sample are fast rotators \citep{atlas3dIII-MNRAS,krajnovic2013}.
%\cite{krajnovic2013} showed that ``fast rotators'' as a class are disk-galaxies or at least disk-like galaxies. 
%In their magnitude-limited survey, systems 
%without any signature of disk-like components (neither in the kinematics nor in the photometry) 
%dominate only the most massive end (beyond $10^{11.5}~\rm M_\odot$) of the distribution. 

@@

and bars can spin-up classical bulges (e.g.~\citealt{saha2012}), 

@@

\subsection{Black hole mass -- spheroid stellar mass}
Finally, we present the $M_{\rm BH} - M_{\rm *,sph}$ diagram in Figure \ref{fig:mbhmasssph}, 
and its linear regression analysis in Table \ref{tab:lregmass}. 
The bulges of the early-type galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$,
consistent with a dry-merging formation scenario\footnote{\bf In dry mergers, the black hole and the bulge grow at the same pace, 
increasing their mass in lock steps.},
and define a tight \emph{early-type sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$. 
\cite{graham2012bent} reported that the $M_{\rm BH}/M_{\rm dyn,sph}$ ratio for core-S\'ersic galaxies was 0.36\% 
($M_{\rm dyn,sph}$ is the spheroid dynamical mass), 
and discussed the many implications of this.  
Using a larger data sample, \cite{grahamscott2013} reported that the $M_{\rm BH}/M_{\rm *,sph}$ ratio was 0.49\% for core-S\'ersic galaxies.   
Here we find a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.50 \pm 0.04\%$ for the 22 core-S\'ersic galaxies 
and $0.68 \pm 0.04\%$ for the 45 early-type galaxies.  
Among other things, this higher value (previously reported to be $0.1 - 0.2\%$ for all galaxy types, e.g.~\citealt{marconihunt2003}), 
boosts estimates of the black hole mass function and mass density based on galaxy/spheroid luminosity functions. \\
The bulges of the spiral galaxies trace a steeper \emph{late-type sequence}, 
whose slope is less well constrained due to the smaller size of the subsample and, more importantly, 
to the smaller range in $M_{\rm *,sph}$ that the subsample spans. 
For the bulges of spiral galaxies, the BCES code returns a log-linear relation with a slope $= 3.00 \pm 1.30$, 
while the modified FITEXY routine finds a shallower (but still consistent within the $1\sigma$ uncertainty) 
slope $= 2.28^{+1.67}_{-1.01}$.
%The Bayesian estimator of \cite{linmixerr} fails in performing an inverse $(X|Y)$ linear regression for the subsample of spiral galaxies. 
More data would be welcome to better constrain the slope of this \emph{late-type sequence}, 
although we note that direct measurements of black hole masses below $10^6\rm~M_\odot$ are extremely challenging to obtain 
with the current technological resources. 
In this regard, using a sample of $\sim$140 low-redshift ($z \leq 0.35$, with a median redshift $\langle z \rangle = 0.085$) 
bulges hosting Active Galactic Nuclei (AGNs) with virial
black hole masses $10^5 \lesssim M_{\rm BH}/{\rm M_\odot} \lesssim 2 \times 10^6$ \citep{jiang2011a}, 
\cite{grahamscott2015} showed that they roughly follow the quadratic $M_{\rm BH} - M_{\rm *,sph}$ relation defined by their S\'ersic bulges.
The majority of our spiral galaxies host an AGN\footnote{According to the nuclear classification reported on NED 
(NASA Extragalactic Database), among our 17 spiral galaxies, at least 12 host a Seyfert AGN and one hosts a LINER AGN.} and
we anticipate here that the correlation traced by our spiral galaxy bulges 
may track the location of these lower mass AGN in the $M_{\rm BH} - M_{\rm *,sph}$ diagram.
That is, the AGNs appear to be the low-mass continuation of our tentative \emph{late-type sequence} shown in Figure \ref{fig:mbhmasssph} 
and this will be explored with more rigour in a forthcoming paper.

>>>>

\title{Supermassive black holes and their host spheroids \\ II. The red and blue sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram}

@@

Kinematical information \citep{atlas3dIII,scott2014,arnold2014} was used 

@@

errors are dominated by systematics rather than Poisson noise. 
{\bf Unlike many papers, we believe that we have not under-estimated the uncertainties associated to the bulge best-fit parameters.} } 

@@

@@

and bars can spin-up classical bulges (e.g.~\citealt{saha2012,saha2015}), 

@@

\subsection{Black hole mass -- spheroid stellar mass}
Finally, we present the $M_{\rm BH} - M_{\rm *,sph}$ diagram in Figure \ref{fig:mbhmasssph}, 
and its linear regression analysis in Table \ref{tab:lregmass}. 
The bulges of the early-type galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$,
consistent with a dry-merging formation scenario\footnote{\bf In dry mergers, the black hole and the bulge grow at the same pace, 
increasing their mass in lock step.},
and define a tight \emph{early-type sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$. 
\cite{graham2012bent} reported that the $M_{\rm BH}/M_{\rm dyn,sph}$ ratio for core-S\'ersic galaxies was 0.36\% 
($M_{\rm dyn,sph}$ is the spheroid dynamical mass), 
and discussed the many implications of this.  
Using a larger data sample, \cite{grahamscott2013} reported that the $M_{\rm BH}/M_{\rm *,sph}$ ratio was 0.49\% for core-S\'ersic galaxies.   
Here we find a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.50 \pm 0.04\%$ for the 22 core-S\'ersic galaxies 
and $0.68 \pm 0.04\%$ for the 45 early-type galaxies.  
Among other things, this higher value (previously reported to be $0.1 - 0.2\%$ for all galaxy types, e.g.~\citealt{marconihunt2003}), 
boosts estimates of the black hole mass function and mass density based on galaxy/spheroid luminosity functions. \\
The bulges of the spiral galaxies trace a steeper \emph{late-type sequence}, 
whose slope is less well constrained due to the smaller size of the subsample and, more importantly, 
to the smaller range in $M_{\rm *,sph}$ that the subsample spans. 
For the bulges of spiral galaxies, the BCES code returns a log-linear relation with a slope $= 3.00 \pm 1.30$, 
while the modified FITEXY routine finds a shallower (but still consistent within the $1\sigma$ uncertainty) 
slope $= 2.28^{+1.67}_{-1.01}$.
%The Bayesian estimator of \cite{linmixerr} fails in performing an inverse $(X|Y)$ linear regression for the subsample of spiral galaxies. 
More data would be welcome to better constrain the slope of this \emph{late-type sequence}, 
although we note that direct measurements of black hole masses below $10^6\rm~M_\odot$ are extremely challenging to obtain 
with the current technological resources. 
In this regard, using a sample of $\sim$140 low-redshift ($z \leq 0.35$, with a median redshift $\langle z \rangle = 0.085$) 
bulges hosting Active Galactic Nuclei (AGNs) with virial
black hole masses $10^5 \lesssim M_{\rm BH}/{\rm M_\odot} \lesssim 2 \times 10^6$ \citep{jiang2011a}, 
\cite{grahamscott2015} showed that they roughly follow the quadratic $M_{\rm BH} - M_{\rm *,sph}$ relation defined by their S\'ersic bulges.
The majority of our spiral galaxies host an AGN\footnote{According to the nuclear classification reported on NED 
(NASA Extragalactic Database), among our 17 spiral galaxies, at least 12 host a Seyfert AGN and one hosts a LINER AGN.} and
we anticipate here that the correlation traced by our spiral galaxy bulges 
may track the location of these lower mass AGN in the $M_{\rm BH} - M_{\rm *,sph}$ diagram.
That is, the AGNs appear to be the low-mass continuation of our tentative \emph{late-type sequence} shown in Figure \ref{fig:mbhmasssph} 
and this will be explored with more rigour in a forthcoming paper.
